Parental Migration and Left-Behind Children's Cognitive and Academic Performances

### 2022-06-03 MScThesisText.Rmd
bring back unadjusted model

---
title             : "Parental Migration and Left-Behind Adolescent's School Performance"
shorttitle        : "Left-Behind Adolescent"

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "j4.gu@student.vu.nl"

@@

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
r_refs("references.bib")
```

@@

Millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called "left behind children" who are common in migrant-sending regions in the Global South. The global size of left behind children population is unavailable, but in-country surveys provide some estimation. [@dewaard2018] analyzed the Mexican and Latin American Migration Projects and estimated that around 15 percent of children were left behind in Mexico, El Salvador, Nicaragua, and Puerto Rico in the 2000s. [@popova2018] cited a 2016 study by UN Children's Fund Bulgaria, suggesting that one-fifth of Bulgarian children had one or both parents working abroad. In China, the 2015 one-percent National Population Sample Survey showed that one out of four children lived under parental absence induced by migration [@nbs2017].

@@

A challenge arises in the empirical assessment of the causal relation between parental migration and children's education. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Statistical advances in causal inference provides an alternative. We can infer causality from a quasi-experimental design -- that is, only if the cause is as good as randomly assigned to individuals in the analysis, conditional on identification assumptions.

@@

@@

I obtained the data from the China Education Panel Survey (CEPS). It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The project's sampling adopted a multi-stage stratified probability proportional to size approach. Moreover, designed deliberately to focus on migrant children, the survey also oversampled those counties with large intranational migrant populations. The participants filled a paper-and-pencil questionnaire. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. I only kept the 9,449 observations of seventh grade students and their parents, who completed the follow-up survey in the 2014-2015 academic year.

@@

I explored two categories of adolescence development: academic and cognitive abilities. One key dependent variable, the student's cognitive skill, is measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The raw scores of cognitive skill were then standardized. The other set of dependent variables concerned students' academic performance. This was measured by the mid-term exam scores in three subjects: Chinese, mathematics, and English. To facilitate comparison across schools and counties, I used standardized test scores instead of raw test scores. I removed the observations with missing values in any of the test score mentioned above.

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I identified seven types of migration arrangement: both parents absent, only father absent, only mother absent, father absent (unconditional), mother absent (unconditional), and both parents at home. I then removed the observations that reported other types of parental absence induced by divorce or death.

@@

Following Bai and colleagues' [-@bai2018; -@bai2020] approach, I adopted a difference-in-differences design to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. The model is: $$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \theta \cdot X_{i,s} + \varepsilon_{i,s}$$For student $i$ in school $s$, $\Delta score_{i,s}$ is the change in standardized test score between baseline and follow-up surveys, $migr_{i,s}$ is the treatment dummy variable, $FE_{s}$ is the school-level fixed effect, $score_{i,s;base}$ is the baseline exam score, and $X_{i,s}$ is the vector of covariates capturing the characteristics of children and their households.

The control variables include the internet assess at home, the self-rated number of books at home, and the years of education of the best-educated parent. These variables can approximate the level of human capital in a household. The internet assess was measured by the item "Do [sic] your family own a computer and have an access to the Internet". Those who answered "Yes, we have both" were coded as 1, and others as 0. The number of books was measured by the item "How many books do your family own? (not including textbooks or magazines)". Students chose from a range of 1 as "very few" to 5 as "a great number". The years of education of the best-educated parent were measured by two items. Students indicated their father's and mother's educational level, respectively. I transformed these two variables into years of education and retained the larger value.

I adopted `r cite_r("r-references.bib")` for all the analyses and the `fixest` package [@bergé2018] for building the two-way fixed effect models.

>>>>

---
title             : "Parental Migration and Left-Behind Children's Cognitive and Academic Performances"
shorttitle        : "Left-Behind Children"

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "De Boelelaan 1105, 1081 HV - Amsterdam, The Netherlands"
    email         : "j4.gu@student.vu.nl"

@@

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
library("modelsummary")
r_refs("references.bib")
```

@@

Millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work [@antia2020]. These are the so-called "left behind children" who are common in migrant-sending regions in the Global South. The global size of left behind children population is unavailable, but in-country surveys provide some estimation. [@dewaard2018] analyzed the Mexican and Latin American Migration Projects and estimated that around 15 percent of children were left behind in Mexico, El Salvador, Nicaragua, and Puerto Rico in the 2000s. [@popova2018] cited a 2016 study by UN Children's Fund Bulgaria indicating that one-fifth of Bulgarian children had one or both parents working abroad. In China, the 2015 one-percent National Population Sample Survey showed that one out of four children lived under parental absence induced by migration [@nbs2017].

@@

A challenge arises in when we want to assess the causality between parental migration and children's education. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Statistical advances in causal inference provides an alternative approach. We can infer causality from a quasi-experimental design -- that is, only if the cause is as good as randomly assigned to individuals in the analysis, conditional on identification assumptions.

@@

I hypothesized that in the short term, parental migration negatively affect the cognitive development and academic performance of children left behind. This is because parental migration immediately disrupts family structure, resulting in lower academic performance of children. It takes longer time for parental migration to generate resources that improve children's performance at school.

Researchers have studied whether the impact differ between mother absence and father absence in the context of labor migration. [@xu2019] adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). [@chen2019] conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

I hypothesized that in the short term, mother migration negatively affect the cognitive development and academic performance of children left behind, while father migration does not show a effect. This relates to the social norms that designate men as breadwinners and women as caregivers. Studies noted that in some Asian and Latin American societies, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left behind children may even resent their migrant mothers, viewing them as transgressing on the expected care-giving roles.

@@

I obtained the data from the China Education Panel Survey (CEPS). It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The research team adopted the approach of multi-stage stratified probability proportional to size sampling. The selected participants filled a paper-and-pencil questionnaire. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. I only kept the 9,449 observations of seventh grade students and their parents, who completed the follow-up survey in the 2014-2015 academic year.

@@

I explored two categories of adolescence development: academic and cognitive abilities. One key dependent variable, the student's cognitive skill, is measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The raw scores of cognitive skill were then standardized. The other set of dependent variables concerned students' academic performance. This was measured by the mid-term exam scores in three subjects: Chinese, mathematics, and English. I calculated the standardized exam grades at the school-by-grade level to facilitate comparison across schools and counties. I removed the observations with missing values in any of the test score mentioned above. My choice of standardized exam grades over raw grades follows previous research [@young2018; @xu2019].

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed six treatment dummy variables measuring migration arrangements: both parents absent, only father absent, only mother absent, father absent (unconditional), mother absent (unconditional). These arrangements can overlap with each other. The control group consisted of parents staying with their children throughout at both the baseline and the follow-up surveys. I then removed the observations that reported other types of parental absence induced by divorce or death.

The control variables at the household level include the internet assess at home, the self-rated number of books at home, and the years of education of the best-educated parent. These variables can approximate the level of human capital in a household. The internet assess was measured by the item "Do [sic] your family own a computer and have an access to the Internet". Those who answered "Yes, we have both" were coded as 1, and others as 0. The number of books was measured by the item "How many books do your family own? (not including textbooks or magazines)". Students chose from a range of 1 as "very few" to 5 as "a great number". The years of education of the best-educated parent were measured by two items. Students indicated their father's and mother's educational level, respectively. I transformed these two variables into years of education and retained the larger value.

@@

I adopted `r cite_r("r-references.bib")` for all the analyses, the `fixest` package for building the econometric models [@bergé2018], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. The model is: $$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \theta \cdot X_{i,s} + \varepsilon_{i,s}$$For student $i$ in school $s$, $\Delta score_{i,s}$ is the change in standardized test score between baseline and follow-up surveys, $migr_{i,s}$ is the treatment dummy variable, $FE_{s}$ is the school-level fixed effect, $score_{i,s;base}$ is the standardized test score at baseline, and $X_{i,s}$ is the vector of covariates capturing the characteristics of children and their households. The standard errors were clustered at the school level.
